---
title: "Data Cleaning for transplant surgeries"
author:
  - Hamidreza Ahady Dolatsara^[Aubrun University, [hamid@auburn.edu](mailto:hamid@auburn.edu)]
  - Fadel M. Megahed^[Miami University, [fmegahed@miamioh.edu](mailto:fmegahed@miamioh.edu)]
  - Ying-Ju Tessa Chen^[University of Dayton, [ychen4@udayton.edu](mailto:ychen4@udayton.edu)]

date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    code_folding: hide
---

  
****

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

**In the First Section I load libraries and import the data**
```{r, load_libraries, message=FALSE, error=FALSE}
# Initilization
#cat("\014") # Clear the console
rm(list=ls())
graphics.off()

require(pacman)
p_load(dplyr,caret,foreign,
       lubridate,dataPreparation,httr, DT,stringr,AUC,parallel, testit,caretEnsemble,
       C50,
       randomForest,
       kernlab,
       e1071,DT)

source("https://raw.githubusercontent.com/hamidahady/transplant/Hamid/isotonic_paper_functions.R")

############################ Data Gathering #########################
# first I download the data from github and then merge them into one large file named heart.df
{
  # file1<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_1_9000L.rds?raw=true"))
  # file2<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_9001_18000L.rds?raw=true"))
  # file3<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_18001_27000L.rds?raw=true"))
  # file4<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_27001_36000L.rds?raw=true"))
  # file5<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_36001_45000L.rds?raw=true"))
  # file6<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_45001_54000L.rds?raw=true"))
  # file7<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_54001_63000L.rds?raw=true"))
  # file8<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_63001_72000L.rds?raw=true"))
  # file9<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_72001_81000L.rds?raw=true"))
  # file10<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_81001_90000L.rds?raw=true"))
  # file11<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_90001_99000L.rds?raw=true"))
  # file12<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_99001_108000L.rds?raw=true"))
  # file13<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_108001_117000L.rds?raw=true"))
  # file14<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_117001_126000L.rds?raw=true"))
  # file15<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_126001_135000L.rds?raw=true"))
  # file16<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_135001_144000L.rds?raw=true"))
  # file17<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_144001_153000L.rds?raw=true"))
  # file18<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_153001_159318L.rds?raw=true"))
  
  file1<-readRDS(("C:/Users/hza0020/Box/Transplant/Thesis/github/transplant/transplant/data/raw/thoracic_data_1_9000L.rds"))
  file2<-readRDS(("C:/Users/hza0020/Box/Transplant/Thesis/github/transplant/transplant/data/raw/thoracic_data_9001_18000L.rds"))
  file3<-readRDS(("C:/Users/hza0020/Box/Transplant/Thesis/github/transplant/transplant/data/raw/thoracic_data_18001_27000L.rds"))
  file4<-readRDS(("C:/Users/hza0020/Box/Transplant/Thesis/github/transplant/transplant/data/raw/thoracic_data_27001_36000L.rds"))
  file5<-readRDS(("C:/Users/hza0020/Box/Transplant/Thesis/github/transplant/transplant/data/raw/thoracic_data_36001_45000L.rds"))
  file6<-readRDS(("C:/Users/hza0020/Box/Transplant/Thesis/github/transplant/transplant/data/raw/thoracic_data_45001_54000L.rds"))
  file7<-readRDS(("C:/Users/hza0020/Box/Transplant/Thesis/github/transplant/transplant/data/raw/thoracic_data_54001_63000L.rds"))
  file8<-readRDS(("C:/Users/hza0020/Box/Transplant/Thesis/github/transplant/transplant/data/raw/thoracic_data_63001_72000L.rds"))
  file9<-readRDS(("C:/Users/hza0020/Box/Transplant/Thesis/github/transplant/transplant/data/raw/thoracic_data_72001_81000L.rds"))
  file10<-readRDS(("C:/Users/hza0020/Box/Transplant/Thesis/github/transplant/transplant/data/raw/thoracic_data_81001_90000L.rds"))
  file11<-readRDS(("C:/Users/hza0020/Box/Transplant/Thesis/github/transplant/transplant/data/raw/thoracic_data_90001_99000L.rds"))
  file12<-readRDS(("C:/Users/hza0020/Box/Transplant/Thesis/github/transplant/transplant/data/raw/thoracic_data_99001_108000L.rds"))
  file13<-readRDS(("C:/Users/hza0020/Box/Transplant/Thesis/github/transplant/transplant/data/raw/thoracic_data_108001_117000L.rds"))
  file14<-readRDS(("C:/Users/hza0020/Box/Transplant/Thesis/github/transplant/transplant/data/raw/thoracic_data_117001_126000L.rds"))
  file15<-readRDS(("C:/Users/hza0020/Box/Transplant/Thesis/github/transplant/transplant/data/raw/thoracic_data_126001_135000L.rds"))
  file16<-readRDS(("C:/Users/hza0020/Box/Transplant/Thesis/github/transplant/transplant/data/raw/thoracic_data_135001_144000L.rds"))
  file17<-readRDS(("C:/Users/hza0020/Box/Transplant/Thesis/github/transplant/transplant/data/raw/thoracic_data_144001_153000L.rds"))
  file18<-readRDS(("C:/Users/hza0020/Box/Transplant/Thesis/github/transplant/transplant/data/raw/thoracic_data_153001_159318L.rds"))
  
  
  heart.df<-rbind(file1,file2,file3,file4,file5,file6,file7,file8,file9,file10,file11,file12,file13,file14,file15,file16,
                  file17,file18)
  
  rm(list = c("file1","file2","file3","file4","file5","file6","file7","file8","file9","file10","file11","file12","file13",
              "file14","file15","file16","file17","file18"))
}
#####end of importing the data

# heart.form contains the variable definitions of each variable in data
heart.form <- read.csv("C:/Users/hza0020/Box/Transplant/Thesis/github/transplant/transplant/data/var_desc.csv")  

```
**In the next Section, we drop some variables and observations, based on these criterias:**
 - variables associated to before the surgery data
 - variables that did not end & those added before 2000
 - variables that are not interesting/relevant for survival analysis (e.g. follow up number, Donor ID, Dates, and ...)
 - patients who are older than 18 years old
 - patients that are associated extreme 0.1 lightest people
 - patients that are associated extreme 0.1 shortest people
 _ patient who did not have info. of transplanted organ
 - patients who has lung transplant, too.
 
```{r}
# Identifying the variables that are discarded for further analysis
vars_discarded <- heart.form %>% # We created a column a colum INTERPRETATION_TYPE  
  subset(INTERPRETATION_TYPE=="D", select=c(1,2))
heart.discard <- vars_discarded$VARIABLE.NAME %>% as.character()

# Identifying the variables that did not end & those added before 2000
var.end.dates <- trimws(heart.form$VAR.END.DATE) %>% str_trim()
if.var.did.not.end <- which(var.end.dates=="")
names.of.var.did.not.end <- heart.form[if.var.did.not.end, 1] 

# Identifying the variables that ended
vars_ended <-heart.form[which(heart.form$VARIABLE.NAME %in% names(heart.df)[!(names(heart.df) %in% names.of.var.did.not.end)]),(1:2)]


vars.added.dates <- heart.form$VAR.START.DATE %>% as.character.Date()
vars.added.dates[which(vars.added.dates=="01-Oct-87, 01-Oct-90")] <- "01-Oct-90"
vars.added.years<- sapply(vars.added.dates, function(x) str_extract_all(x,"[0-9]{1,2}")[[1]][2]) %>% as.integer()
heart.form$YR_ADDED <- vars.added.years
vars.added.before.2000 <- subset(heart.form,YR_ADDED>=87,
                                 select = c(1))  # available variables added before 2000
vars.added.NA <- subset(heart.form,is.na(YR_ADDED),select = c(1))
vars.added.all <- rbind(vars.added.before.2000,vars.added.NA)
vars.added.all <- vars.added.all[["VARIABLE.NAME"]] %>% 
  as.character()

# Subsetting the data nd getting rid of variables that are ended and extreme patients or patients that are less than 18
heart.df.cleaned <- subset(heart.df, WL_ORG=="HR") %>% # Heart 
  subset(AGE>=18) %>% # Adults only
  # we excluded too light or too short people
  subset(WGT_KG_DON_CALC >= quantile(WGT_KG_DON_CALC, 0.0001, na.rm = TRUE)) %>% 
  subset(WGT_KG_TCR >= quantile(WGT_KG_TCR, 0.0001, na.rm = TRUE)) %>%
  subset(HGT_CM_DON_CALC >= quantile(HGT_CM_DON_CALC, 0.0001, na.rm = TRUE)) %>% 
  subset(HGT_CM_TCR >= quantile(HGT_CM_TCR, 0.0001, na.rm = TRUE)) %>% 
  subset(select=intersect(names.of.var.did.not.end,vars.added.all))


#Identifying and removing the variables that are post transplant
# Based on observations from the Form and Form Section Descriptors
vars.post.trans.index1 <-  sapply(heart.form$FORM.SECTION,
                                  function(x) str_detect(x,
                                                         "POST TRANSPLANT CLINICAL INFORMATION"))
vars.post.trans.index2 <-  sapply(heart.form$FORM,
                                  function(x) str_detect(x,
                                                         "TRF/TRR|TRR/TRF-CALCULATED|TRR/TRF|TRF"))
vars_post<- heart.form[as.logical(vars.post.trans.index1+vars.post.trans.index2),][,1] %>%
  as.character()

# Identifying the post transplant variables
vars_posttrans<-heart.form[which(heart.form$VARIABLE.NAME %in% vars_post),(1:2)]

vars.post.trans <- intersect(colnames(heart.df.cleaned),
                             vars_post)
heart.df.cleaned <- select(heart.df.cleaned,-vars.post.trans)
heart.discard <- intersect(heart.discard, colnames(heart.df.cleaned))
heart.df.cleaned <- select(heart.df.cleaned, -heart.discard) 

```
<br />
<br />
**Here are the variables that we dropped for bring not interesting/irrelavence**
```{r , message=FALSE, cache=FALSE, error=FALSE}
DT::datatable(vars_discarded)
```
<br />
<br />
**Here are the variables that ended. Their adding date is chaecked, even if they dropped late (2015) but the added early so they are intersting (e.g. PRAMR_CL2).**
```{r , message=FALSE, cache=FALSE, error=FALSE}
DT::datatable(vars_ended)
```
<br />
<br />
**Here are the variables that are relevent to the post transplant info.**
```{r , message=FALSE, cache=FALSE, error=FALSE}
DT::datatable(vars_posttrans)
```
<br />
<br />
**Here are a report about the remaining variables (discarded means irrelevant/not interesting variables)**
**We lready got rid of discarded variables so their frequency is zero**
```{r , message=FALSE, cache=FALSE, error=FALSE}

rem_type<-as.data.frame(table(heart.form[which(heart.form$VARIABLE.NAME %in% names(heart.df.cleaned)), "INTERPRETATION_TYPE"]))
names(rem_type)<-c("Variable Type","Frequency")
rem_type$`Variable Type`<-c("Categorical","Discarded","Date","Numerical")

org_type<-as.data.frame(table(heart.df$WL_ORG))
names(org_type)<-c("Organ Type","No. of Patients")
org_type$`Variable Type`<-c("UNKOWN","Heart & Lung","Heart","Lung")
DT::datatable(org_type)
DT::datatable(rem_type)

cat("Number of patients after dropping irrelevant patients: ",nrow(heart.df.cleaned))

```
<br />
<br />
<br />
<br />
**inthe following section variables are recategorized, drived, and developed based on literature considering the pool of patients**
```{r , message=FALSE, cache=FALSE, error=FALSE}
# ref1: Medved, Dennis, et al. "Improving prediction of heart transplantation outcome using deep learning techniques." Scientific reports 8.1 (2018): 3613.
  # refer to a tool provided in the: http://ihtsa.cs.lth.se/ , which is product of this paper:
  # https://www.nature.com/articles/s41598-018-21417-7.pdf
  
  # In this paper they used these variables for recipients:
  # Diagnosis, Age, Gender, Height, Weight, insulin treated diabetes, infection within two weeks, Blood group, previous blood transfusion,
  # previously transplanted, previous cardiac surgery, intensive care unit, mechanical ventilation,
  # ECMO, ventiricular assist device, transplant era, SPP, PVR, creatinine, serum bilirubin,
  # use (mg/dl) instead of (mio mol/l), PRA>10%, HLA-DR 2 mismatch
  # Also, they used these variables for donor:
  # Age, Gender, Height, Weight, Duration of ischemia, Blood group, cause of death

  # we also used the following variables without any change:
  # Age / AGE, AGE_DON
  # Height / HGT_CM_CALC, HGT_CM_DON_CALC, HGT_CM_TCR, INIT_HGT_CM_CALC, 
  # Gender / GENDER, GENDER_DON
  # Weight / WGT_KG_DON_CALC, WGT_KG_TCR, WGT_KG_CALC, PERCENT_WGT_CHANGE (we made it)
  # infection within two weeks / INFECT_IV_DRUG_TRR
  # previous blood transfusion / TRANSFUSIONS
  # Previously transplanted / PREV_TX
  # Intensive care unit / ICU
  # ventiricular assist device / VAS
  # Mechanical ventilation / VENTILATOR_TRR
  # SPP (mmHG), systolic pulmonary pressure  / HEMO_SYS_TRR
  # Creatinine / CREAT_TRR
  # Serum bilirubin (μmol/l) / TBILI
  # Use (mg/dl) instead of (μmol/l) / TBILI
  # we used mg/dl
  # we did not work with PRAMR since it is added after 2004

# ref2: Dag, Ali, et al. "Predicting heart transplantation outcomes through data analytics." Decision Support Systems 94 (2017): 42-52.

# Subsetting the data and creating other variables based on literature or to capture time effects
base<-heart.df.cleaned
heart.df.cleaned<-base
table(heart.df.cleaned$HEMO_PA_MN_TRR)

heart.df.cleaned <- subset(heart.df.cleaned) %>%  
  #ref1
  mutate(PVR = (HEMO_PA_MN_TRR- HEMO_PCW_TRR)*79.72/HEMO_CO_TRR) %>% 
  #ref1
  mutate(ISCHTIME = ISCHTIME*60) %>% 
  #ref1
  mutate(ECMO = ifelse(ECMO_TCR + ECMO_TRR == 0, 0, 1)) %>% 
  # PVR, pulmonary vascular resistance / its calculation is based on the below mentioned links:
  # https://en.wikipedia.org/wiki/Vascular_resistance
  # http://www.scymed.com/en/smnxph/phkhr013.htm
  # https://radiopaedia.org/articles/mean-pulmonary-arterial-pressure  (calculation of Mean Pulmonary Arterial Pressure)
  # PVR= (Mean Pulmonary Arterial Pressure (mmHg) - Pulmonary Capillary Wedge Pressure (mmHg)) * 79.72 / Cardiac Output (L/min)
  # PVR = (HEMO_PA_MN_TRR - HEMO_PCW_TRR)* 79.72 / HEMO_CO_TRR

  # ECMO / merge of (ECMO_TCR, ECMO_TRR)

  # The following variables are mutated by the author
  mutate(BMI_CHNG = 100*(BMI_CALC- INIT_BMI_CALC)/INIT_BMI_CALC) %>%
  mutate(WAITING_TIME = TX_DATE - INIT_DATE)  %>%
  mutate(WGT_CHNG = 100*(WGT_KG_CALC - INIT_WGT_KG_CALC)/INIT_WGT_KG_CALC) %>%
  mutate(HGT_CHNG = 100*(HGT_CM_CALC - INIT_HGT_CM_CALC)/INIT_HGT_CM_CALC) %>%
  mutate(AGE_MAT = abs(AGE - AGE_DON)) %>%
  mutate(BMI_MAT = abs(BMI_CALC - BMI_DON_CALC))

# we did below variable based on the previous section
base[c("HEMO_PA_MN_TRR" , "HEMO_PCW_TRR","HEMO_CO_TRR","ECMO_TCR" , "ECMO_TRR","BMI_CALC", "INIT_BMI_CALC",
       "TX_DATE" , "INIT_DATE", "WGT_KG_CALC" , "INIT_WGT_KG_CALC", "HGT_CM_CALC" , "INIT_HGT_CM_CALC",
       "AGE" , "AGE_DON" , "BMI_DON_CALC")]<-NULL
# we do not need anything on the beow variable based on ref1
 base[ c("AGE", "AGE_DON","HGT_CM_CALC", "HGT_CM_DON_CALC", "HGT_CM_TCR", "INIT_HGT_CM_CALC",
  "GENDER", "GENDER_DON", "WGT_KG_DON_CALC", "WGT_KG_TCR", "WGT_KG_CALC", "INFECT_IV_DRUG_TRR",
  "TRANSFUSIONS", "PREV_TX", "ICU", "VAS", "VENTILATOR_TRR","HEMO_SYS_TRR", "CREAT_TRR","TBILI")]<-NULL

 # DIAG was used in ref1, but it has same form with TCR_DGN and THORACIC_DGN so we recategorize all three together
 
table(heart.df.cleaned$TCR_DGN)
table(heart.df.cleaned$DIAG)
table(heart.df.cleaned$THORACIC_DGN)

  heart.df.cleaned$temp<-heart.df.cleaned$DIAG
  for(i in 1:nrow(heart.df.cleaned)){
    if(!is.na(heart.df.cleaned$DIAG[i])){
      heart.df.cleaned$DIAG[i]<-"other"
      if(heart.df.cleaned$temp[i]==1000){heart.df.cleaned$DIAG[i]<-"DILATED_MYOPATHY_IDI"}
      
      if((heart.df.cleaned$temp[i]>=1001 & heart.df.cleaned$temp[i]<= 1006) |
         (heart.df.cleaned$temp[i]==1049)){heart.df.cleaned$DIAG[i]<-"DILATED_MYOPATHY_OTH"}
      
      if(heart.df.cleaned$temp[i]==1007){heart.df.cleaned$DIAG[i]<-"DILATED_MYOPATHY_ISC"}
      
      if(heart.df.cleaned$temp[i]==1200){heart.df.cleaned$DIAG[i]<-"CORONARY"}

    }
  }



sum(as.data.frame(table(heart.df.cleaned$TCR_DGN))[2])
sum(as.data.frame(table(heart.df.cleaned$DIAG))[2])
sum(as.data.frame(table(heart.df.cleaned$THORACIC_DGN))[2])

```







