---
title: "Data Cleaning for transplant surgeries"
author:
  - Hamidreza Ahady Dolatsara^[Aubrun University, [hamid@auburn.edu](mailto:hamid@auburn.edu)]
  - Fadel M. Megahed^[Miami University, [fmegahed@miamioh.edu](mailto:fmegahed@miamioh.edu)]
  - Ying-Ju Tessa Chen^[University of Dayton, [ychen4@udayton.edu](mailto:ychen4@udayton.edu)]

date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    code_folding: hide
---

  
****

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

**In the First Section I load libraries and import the data**
```{r, load_libraries, message=FALSE, error=FALSE}
# Initilization
#cat("\014") # Clear the console
rm(list=ls())
graphics.off()

require(pacman)
p_load(dplyr,caret,foreign,
       lubridate,dataPreparation,httr, DT,stringr,AUC,parallel, testit,caretEnsemble,
       C50,
       randomForest,
       kernlab,
       e1071,DT)

source("https://raw.githubusercontent.com/hamidahady/transplant/Hamid/isotonic_paper_functions.R")

############################ Data Gathering #########################
# first I download the data from github and then merge them into one large file named heart.df
{
  # file1<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_1_9000L.rds?raw=true"))
  # file2<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_9001_18000L.rds?raw=true"))
  # file3<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_18001_27000L.rds?raw=true"))
  # file4<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_27001_36000L.rds?raw=true"))
  # file5<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_36001_45000L.rds?raw=true"))
  # file6<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_45001_54000L.rds?raw=true"))
  # file7<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_54001_63000L.rds?raw=true"))
  # file8<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_63001_72000L.rds?raw=true"))
  # file9<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_72001_81000L.rds?raw=true"))
  # file10<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_81001_90000L.rds?raw=true"))
  # file11<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_90001_99000L.rds?raw=true"))
  # file12<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_99001_108000L.rds?raw=true"))
  # file13<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_108001_117000L.rds?raw=true"))
  # file14<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_117001_126000L.rds?raw=true"))
  # file15<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_126001_135000L.rds?raw=true"))
  # file16<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_135001_144000L.rds?raw=true"))
  # file17<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_144001_153000L.rds?raw=true"))
  # file18<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_153001_159318L.rds?raw=true"))
  

}
#####end of importing the data

# heart.form contains the variable definitions of each variable in data
heart.form <- read.csv("https://raw.githubusercontent.com/hamidahady/transplant/Hamid/data/var_desc.csv")  

```
> ### In the next Section, we drop some variables and observations, based on these criterias:

* variables associated to before the surgery data
* variables that did not end & those added before 2000
* variables that are not interesting/relevant for survival analysis (e.g. follow up number, Donor ID, Dates, and ...)
* patients who are older than 18 years old
* patients that are associated extreme 0.1 lightest people
* patients that are associated extreme 0.1 shortest people
* patient who did not have info. of transplanted organ
* patients who has lung transplant, too.
 