---
title: Investigating the thoracic data
author:

- Hamidreza Ahady Dolatsara^[Auburn University, [hamid@auburn.edu](mailto:hamid@auburn.edu)]
- Fadel Megahed^[University of Miami Ohio, [fmegahed@miamioh.edu](mailto:fmegahed@miamioh.edu)]
- Ying-Ju Tessa Chen^[Dayton University, [ychen4@udayton.edu](mailto:ychen4@udayton.edu)]
- Christy Evans^[Auburn University, [cje0010@auburn.edu](mailto:cje0010@auburn.edu)]
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    code_folding: hide
    df_print: paged
    number_sections: yes
    toc: yes
    toc_float: yes
  word_document:
    toc: yes
---

  

****



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In the following block, required ***R*** libraries are checked for existance and then installed if it's needed.


```{r, message=FALSE, cache=FALSE, error=FALSE}

rm(list = ls()) # clears global environment (functions and variables)
graphics.off() # clears graphics
if(!("pacman" %in% rownames(installed.packages()))){install.packages("pacman")}# pacman would be installed if it's not existed
library(pacman) 
p_load(caret,httr, DT,stringr)

# I load functions from this source
#replace by github
#source("https://raw.githubusercontent.com/hamidahady/transplant/Hamid/all_functions.R")
source("C:/Users/hza0020/Box/Transplant/Thesis/github/transplant/transplant/isotonic_paper_functions.R")

 
```
Here we check different values that a variable might have.
```{r , message=FALSE, cache=TRUE, error=FALSE, results="hide", warning=FALSE}

# first I download the data from github and then merge them into one large file named thoracic_data
{
file1<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_1_9000L.rds?raw=true"))
file2<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_9001_18000L.rds?raw=true"))
file3<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_18001_27000L.rds?raw=true"))
file4<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_27001_36000L.rds?raw=true"))
file5<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_36001_45000L.rds?raw=true"))
file6<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_45001_54000L.rds?raw=true"))
file7<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_54001_63000L.rds?raw=true"))
file8<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_63001_72000L.rds?raw=true"))
file9<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_72001_81000L.rds?raw=true"))
file10<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_81001_90000L.rds?raw=true"))
file11<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_90001_99000L.rds?raw=true"))
file12<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_99001_108000L.rds?raw=true"))
file13<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_108001_117000L.rds?raw=true"))
file14<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_117001_126000L.rds?raw=true"))
file15<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_126001_135000L.rds?raw=true"))
file16<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_135001_144000L.rds?raw=true"))
file17<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_144001_153000L.rds?raw=true"))
file18<-readRDS(url("https://github.com/hamidahady/transplant/blob/master/data/raw/thoracic_data_153001_159318L.rds?raw=true"))


thoracic_data<-rbind(file1,file2,file3,file4,file5,file6,file7,file8,file9,file10,file11,file12,file13,file14,file15,file16,
                     file17,file18)

rm(list = c("file1","file2","file3","file4","file5","file6","file7","file8","file9","file10","file11","file12","file13",
            "file14","file15","file16","file17","file18"))
}
#####end of importing the data
  
var_table<-matrix(NA, ncol = 2, nrow = ncol(thoracic_data))
var_table<-as.data.frame(var_table)
names(var_table) <- c("vars","data")

var_table$vars<-colnames(thoracic_data)

for(i in 1:ncol(thoracic_data)){
  if(length(as.data.frame(table(thoracic_data[i]))[,1])<100){
    var_table$data[i]<-paste(as.data.frame(table(thoracic_data[i]))[,1],collapse=",")
  }else{var_table$data[i]<-"numeric or too many levels"}
  
  
}

# our dataset has patient data from 1987 to 2016
#variables changed overtime, some of them are added very late, it means for majority of patiesnts, no values are recorded.
#Also some variables are taken out of the dataset longtime ago, it means for majority of patiesnts, no values are recorded.

# Step 1: Identifying the variables that did not end & those added before 2000

# first I import a major file about the variables

# in this dataset, the NA values are specified in different formats: NA, "", " ", UNK, ...
# in this block I convert those values to NA for a better investigation
{
var_desc<-read.csv("https://raw.githubusercontent.com/hamidahady/transplant/Hamid/data/var_desc.csv")
thoracic_data<-na_maker(thoracic_data,var_desc$VARIABLE.NAME,var_desc$nas) 
  }
###
#save here
#trimming out their space from date
var.end.dates <- trimws(var_desc$VAR.END.DATE) %>% stringr::str_trim()
if.var.did.not.end <- which(var.end.dates=="")

names.of.var.did.not.end <- var_desc[if.var.did.not.end, "VARIABLE.NAME"] 
  
vars.added.dates <- var_desc$VAR.START.DATE %>% as.character.Date()

#changing variable adding date from "01-Oct-87, 01-Oct-90") to "01-Oct-90"
vars.added.dates[which(vars.added.dates=="01-Oct-87, 01-Oct-90")] <- "01-Oct-90"
vars.added.years<- sapply(vars.added.dates, function(x) str_extract_all(x,"[0-9]{1,2}")[[1]][2]) %>% as.integer()
var_desc$YR_ADDED <- vars.added.years
vars.added.df <- subset(var_desc,(YR_ADDED>=87 | is.na(YR_ADDED)) ,
                                select = c(1))  # available variables added before 2000 or has unkown adding date

#making variables as a vector
vars.added.all <- vars.added.df[["VARIABLE.NAME"]] %>% 
                  as.character()

# we excluded patients who just had heart transplant, and did it after year 2000
# we excluded patients who had heart transplant before 2007, so we have at least 10 years of data
# table(thoracic_data$PRAMR)
# temp<-thoracic_data[which(thoracic_data$TX_YEAR>2007),]
# table(temp$PRAMR)
thoracic_data$ID<-rownames(thoracic_data)
heart.df.cleaned <- subset(thoracic_data,WL_ORG=="HR") %>% 
 # the next line is excluded since they might live more than 12 years but the status is NA, also same for less than 2007
 # some of them might lived for 5 years and died but excluded from here
 # also I did not excluded age since it might be just an important variable, however I excluded excluded too short and too light people
 # since it might be just wrong data or extreme peopl
 # subset(is.na(GTIME)==FALSE)  %>% 
  #subset(TX_YEAR>=2000)  %>%
  # subset(TX_YEAR<=2007)  %>%
  subset(WGT_KG_DON_CALC>=25)  %>% #excluding too light people,
  subset(WGT_KG_TCR>=25)  %>% #excluding too light people,
  subset(INIT_WGT_KG_CALC>=25)  %>%  #excluding too light people,

  # subset(AGE_DON>=18)  %>%  #excluding kids,
  # subset(AGE>=18)  %>% #excluding kids,
  
  subset(HGT_CM_DON_CALC>=18)  %>%  #excluding too sohrt,
  subset(HGT_CM_TCR>=18)  %>% #excluding too short,
  
  subset(select=intersect(names.of.var.did.not.end,vars.added.all))

# I added below because it was used in the literature Nature or others
heart.df.cleaned$PRAMR<-thoracic_data$PRAMR[which(thoracic_data$ID %in% heart.df.cleaned$ID)]
# the below variables are selected from literature (Ali dag)
heart.df.cleaned$FUNC_STAT_TCR<-thoracic_data$FUNC_STAT_TCR[which(thoracic_data$ID %in% heart.df.cleaned$ID)]
heart.df.cleaned$FUNC_STAT_TRR<-thoracic_data$FUNC_STAT_TRR[which(thoracic_data$ID %in% heart.df.cleaned$ID)]


# table(thoracic_data$PRIOR_CARD_SURG_TCR)
# table(thoracic_data$PRIOR_CARD_SURG_TRR)

# end of identifying NAs

# Step 2: Identifying the variables that are post transplant
# Based on observations from the Form and Form Section Descriptors
vars.post.trans.index1 <-  sapply(var_desc$FORM.SECTION,
                                  function(x) str_detect(x,
                                                        "POST TRANSPLANT CLINICAL INFORMATION"))
vars.post.trans.index2 <-  sapply(var_desc$FORM,
                                  function(x) str_detect(x,
                                                         "TRF/TRR|TRR/TRF-CALCULATED|TRR/TRF|TRF"))

vars.post.trans <- var_desc[as.logical(vars.post.trans.index1+vars.post.trans.index2),][,1] %>%
  as.character()

vars.post.trans <- intersect(colnames(heart.df.cleaned),
                             vars.post.trans)

heart.df.cleaned<-heart.df.cleaned[which(!colnames(heart.df.cleaned) %in% vars.post.trans)]

########################

#excluding some irrelevant variables
for(i in 1:ncol(heart.df.cleaned)){
  for (j in 1:nrow(var_desc)) {
    if(names(heart.df.cleaned)[i]==var_desc$VARIABLE.NAME[j]){
      if(var_desc$Hamid_TYPE[j]=="D"){heart.df.cleaned[i]<-NA
      }
    }
  }
}

# droppping columns that are completely NAs
heart.df.cleaned <- heart.df.cleaned[,colSums(is.na(heart.df.cleaned))<nrow(heart.df.cleaned)]
heart.df.cleaned$TX_DATE<-as.character(heart.df.cleaned$TX_DATE)
heart.df.cleaned$INIT_DATE<-as.character(heart.df.cleaned$INIT_DATE)

# change NA equivalents to NA
{
  NA_cells<-c(""," ","U")
  for(i in 1:length(NA_cells)){
    
    heart.df.cleaned[heart.df.cleaned == NA_cells[i]] <- NA
    gc()}
}

# refer to a tool provided in the: http://ihtsa.cs.lth.se/ , which is product of this paper:
# https://www.nature.com/articles/s41598-018-21417-7.pdf

# # in this paper they used these variables for recipients:
# Diagnosis, Age, Gender, Height, Weight, insulin treated diabetes, infection within two weeks, Blood group, previous blood transfusion,
# previously transplanted, previous cardiac surgery, intensive care unit, mechanical ventilation,
# ECMO, ventiricular assist device, transplant era, SPP, PVR, creatinine, serum bilirubin,
# use (mg/dl) instead of (mio mol/l), PRA>10%, HLA-DR 2 mismatch
# # in this paper they used these variables for donor:
# Age, Gender, Height, Weight, Duration of ischemia, Blood group, cause of death

# Age / AGE, AGE_DON
# Height / HGT_CM_CALC, HGT_CM_DON_CALC
# Gender / GENDER, GENDER_DON
# Weight / WGT_KG_DON_CALC, WGT_KG_TCR


# we regroup the Diagnosis as below:

# Diagnosis / DIAG, TCR_DGN, THORACIC_DGN:
# Non-ischemic cardiomyopathy: 1201
# Ischemic: 1007
# Valv heart disease: 1202
# Congenital: 1203 1205 1206 1207
# Other : The rest

heart.df.cleaned$temp<-heart.df.cleaned$DIAG
for(i in 1:nrow(heart.df.cleaned)){
  if(!is.na(heart.df.cleaned$DIAG[i])){
    heart.df.cleaned$DIAG[i]<-"other"
    if(heart.df.cleaned$temp[i]==1201){heart.df.cleaned$DIAG[i]<-"nonisch"}
    if(heart.df.cleaned$temp[i]==1007){heart.df.cleaned$DIAG[i]<-"isch"}
    if(heart.df.cleaned$temp[i]==1202){heart.df.cleaned$DIAG[i]<-"valv"}
    if(heart.df.cleaned$temp[i]==1203){heart.df.cleaned$DIAG[i]<-"cong"}
    if(heart.df.cleaned$temp[i]==1205){heart.df.cleaned$DIAG[i]<-"cong"}
    if(heart.df.cleaned$temp[i]==1206){heart.df.cleaned$DIAG[i]<-"cong"}
    if(heart.df.cleaned$temp[i]==1207){heart.df.cleaned$DIAG[i]<-"cong"}
    
  }
}
table(heart.df.cleaned$DIAG)
# cause of death / COD_CAD_DON:
# Head trauma: 3
# cerebrovascular event: 2
# other: the rest

heart.df.cleaned$temp<-heart.df.cleaned$COD_CAD_DON
for(i in 1:nrow(heart.df.cleaned)){
  if(!is.na(heart.df.cleaned$COD_CAD_DON[i])){
    heart.df.cleaned$COD_CAD_DON[i]<-"rest"
    if(heart.df.cleaned$temp[i]==3){heart.df.cleaned$COD_CAD_DON[i]<-"head"}
    if(heart.df.cleaned$temp[i]==2){heart.df.cleaned$COD_CAD_DON[i]<-"cereb"}
  }
}
table(heart.df.cleaned$COD_CAD_DON)

#here save2
# Blood group / ABO, ABO_DON: 
# A: A1, A2, A
# B: B
# AB: A1B, A2B, AB
# O: O

heart.df.cleaned$temp<-heart.df.cleaned$ABO
for(i in 1:nrow(heart.df.cleaned)){
  if(!is.na(heart.df.cleaned$ABO[i])){
    heart.df.cleaned$ABO[i]<-"AB"
    if(heart.df.cleaned$temp[i]=="A1"){heart.df.cleaned$ABO[i]<-"A"}
    if(heart.df.cleaned$temp[i]=="A2"){heart.df.cleaned$ABO[i]<-"A"}
    if(heart.df.cleaned$temp[i]=="A"){heart.df.cleaned$ABO[i]<-"A"}
    if(heart.df.cleaned$temp[i]=="B"){heart.df.cleaned$ABO[i]<-"B"}
    if(heart.df.cleaned$temp[i]=="O"){heart.df.cleaned$ABO[i]<-"O"}
  }
}
table(heart.df.cleaned$ABO)
heart.df.cleaned$temp<-heart.df.cleaned$ABO_DON
for(i in 1:nrow(heart.df.cleaned)){
  if(!is.na(heart.df.cleaned$ABO_DON[i])){
    heart.df.cleaned$ABO_DON[i]<-"AB"
    if(heart.df.cleaned$temp[i]=="A1"){heart.df.cleaned$ABO_DON[i]<-"A"}
    if(heart.df.cleaned$temp[i]=="A2"){heart.df.cleaned$ABO_DON[i]<-"A"}
    if(heart.df.cleaned$temp[i]=="A"){heart.df.cleaned$ABO_DON[i]<-"A"}
    if(heart.df.cleaned$temp[i]=="B"){heart.df.cleaned$ABO_DON[i]<-"B"}
    if(heart.df.cleaned$temp[i]=="O"){heart.df.cleaned$ABO_DON[i]<-"O"}
  }
}
table(heart.df.cleaned$ABO_DON)
# insulin treated diabetes / DIAB: Refer to this paper, type 1 and type 2 are considered as insulin dependent patient
# Kaufman, D. B., & Sutherland, D. E. (2011). Simultaneous pancreas-kidney transplants are appropriate in insulin-treated candidates 
# with uremia regardless of diabetes type. Clinical Journal of the American Society of Nephrology, 6(5), 957-959.

# Yes: 2,3
# No: 1,4
 
heart.df.cleaned$temp<-heart.df.cleaned$DIAB
for(i in 1:nrow(heart.df.cleaned)){
  if(!is.na(heart.df.cleaned$DIAB[i])){
    heart.df.cleaned$DIAB[i]<-"N"
    if(heart.df.cleaned$temp[i]==2){heart.df.cleaned$DIAB[i]<-"Y"}
    if(heart.df.cleaned$temp[i]==3){heart.df.cleaned$DIAB[i]<-"Y"}
  }
}

table(heart.df.cleaned$DIAB)

# infection within two weeks / INFECT_IV_DRUG_TRR:

# previous blood transfusion / TRANSFUSIONS

# Previously transplanted / PREV_TX

# previous cardiac surgery /merge of (PRIOR_CARD_SURG_TCR, PRIOR_CARD_SURG_TRR)

heart.df.cleaned$CARD_SURG<-NA
for(i in 1:nrow(heart.df.cleaned)){
  if(!is.na(heart.df.cleaned$PRIOR_CARD_SURG_TCR[i])){
  if(heart.df.cleaned$PRIOR_CARD_SURG_TCR[i]=="Y"){heart.df.cleaned$CARD_SURG[i]<-"Y"}
    
  if(heart.df.cleaned$PRIOR_CARD_SURG_TCR[i]=="N"){
    
    if(!is.na(heart.df.cleaned$PRIOR_CARD_SURG_TRR[i])){
    
      if(heart.df.cleaned$PRIOR_CARD_SURG_TRR[i]=="N"){heart.df.cleaned$CARD_SURG[i]<-"N"}
    
    }
    
    
    }
    
    
    
    }
  
  if(!is.na(heart.df.cleaned$PRIOR_CARD_SURG_TRR[i])){
  if(heart.df.cleaned$PRIOR_CARD_SURG_TRR[i]=="Y"){heart.df.cleaned$CARD_SURG[i]<-"Y"}}
  
  
  

}

table(heart.df.cleaned$CARD_SURG)
# which is: TCR PRIOR CARDIAC SURGERY AT LISTING, TRR CARDIAC SURGERY BETWEEN LISTING AND TRANSPLANT
# Intensive care unit / ICU
# ventiricular assist device / VAS
# Mechanical ventilation / VENTILATOR_TRR
# ECMO / merge of (ECMO_TCR, ECMO_TRR)
heart.df.cleaned$ECMO<-NA
for(i in 1:nrow(heart.df.cleaned)){
  if(!is.na(heart.df.cleaned$ECMO_TCR[i]) && !is.na(heart.df.cleaned$ECMO_TRR[i])){
  if(heart.df.cleaned$ECMO_TCR[i] + heart.df.cleaned$ECMO_TRR[i] ==1){heart.df.cleaned$ECMO[i]<-"Y"}
  if(heart.df.cleaned$ECMO_TCR[i] + heart.df.cleaned$ECMO_TRR[i] ==0){heart.df.cleaned$ECMO[i]<-"N"}  
  }}


# SPP (mmHG), systolic pulmonary pressure  / HEMO_SYS_TRR

# PVR, pulmonary vascular resistance / its calculation is based on the below mentioned links:
# https://en.wikipedia.org/wiki/Vascular_resistance
# http://www.scymed.com/en/smnxph/phkhr013.htm
# https://radiopaedia.org/articles/mean-pulmonary-arterial-pressure  (calculation of Mean Pulmonary Arterial Pressure)
# PVR= (Mean Pulmonary Arterial Pressure (mmHg) - Pulmonary Capillary Wedge Pressure (mmHg)) * 79.72 / Cardiac Output (L/min)
# PVR = (HEMO_PA_MN_TRR - HEMO_PCW_TRR)* 79.72 / HEMO_CO_TRR


heart.df.cleaned$PVR<-NA
for(i in 1:nrow(heart.df.cleaned)){
  if(!is.na(heart.df.cleaned$HEMO_PA_MN_TRR[i])){
  if(!is.na(heart.df.cleaned$HEMO_PCW_TRR[i])){
  if(!is.na(heart.df.cleaned$HEMO_CO_TRR[i])){
           heart.df.cleaned$PVR[i]<- 79.72*(heart.df.cleaned$HEMO_PA_MN_TRR[i]-heart.df.cleaned$HEMO_PCW_TRR[i])/heart.df.cleaned$HEMO_CO_TRR[i]
  }}}}


# Duration of ischemia (min) / ISCHTIME * 60

for(i in 1:nrow(heart.df.cleaned)){
  if(!is.na(heart.df.cleaned$ISCHTIME[i])){ heart.df.cleaned$ISCHTIME[i]<-heart.df.cleaned$ISCHTIME[i]*60}}
    
 
# Creatinine / CREAT_TRR



# Serum bilirubin (μmol/l) / TBILI

# Use (mg/dl) instead of (μmol/l) / TBILI
# we used mg/dl

# PRA > 10 % / PRAMR
# if PRAMR> 10 Y, otherwise: N


heart.df.cleaned$PRAMR_CAT10<-NA
for(i in 1:nrow(heart.df.cleaned)){
  if(!is.na(heart.df.cleaned$PRAMR[i])){ 
    if(heart.df.cleaned$PRAMR[i]>10){heart.df.cleaned$PRAMR_CAT10[i]<-"Y"}
    if(heart.df.cleaned$PRAMR[i]<=10){heart.df.cleaned$PRAMR_CAT10[i]<-"N"}
        }
  }
table(heart.df.cleaned$PRAMR_CAT10)
# Here save3
# HLA-DR 2 mismatch / RDR2:
# any values: Y, otherwise: 0

for(i in 1:nrow(heart.df.cleaned)){
  if(!is.na(heart.df.cleaned$RDR2[i])){ 
           heart.df.cleaned$RDR2[i]<-"Y"
        }else{heart.df.cleaned$RDR2[i]<-"N"}
}


sum(as.data.frame(table(heart.df.cleaned$RDR2))[2])
#here save4
# Transplant era / TX_YEAR:
# I focused on after 2000. however, I can do for the rest, together or 4 different modeling
# -1995
# 1996-2000
# 2001-2005
# 2006-
heart.df.cleaned$TX_YEAR_CAT<-heart.df.cleaned$TX_YEAR

# categoirizing the transplantation date
for(i in 1:nrow(heart.df.cleaned)){
  if(!is.na(heart.df.cleaned$TX_YEAR[i])){ 
    if(heart.df.cleaned$TX_YEAR[i]<=1995){heart.df.cleaned$TX_YEAR_CAT[i]<-"A"}
    else if(heart.df.cleaned$TX_YEAR[i]>1995 && heart.df.cleaned$TX_YEAR[i]<=2000){heart.df.cleaned$TX_YEAR_CAT[i]<-"B"}
    else if(heart.df.cleaned$TX_YEAR[i]>2000 && heart.df.cleaned$TX_YEAR[i]<=2005){heart.df.cleaned$TX_YEAR_CAT[i]<-"C"}
    else if(heart.df.cleaned$TX_YEAR[i]>2005){heart.df.cleaned$TX_YEAR_CAT[i]<-"D"}
        }
}


#Here save4

# finding duration of waiting time for the operation 

for(i in 1:nrow(heart.df.cleaned)){

  if(!is.na(heart.df.cleaned$INIT_DATE[i]) && !is.na(heart.df.cleaned$TX_DATE[i])){ 
    heart.df.cleaned$WAITING_TIME[i]<-as.numeric(as.Date(heart.df.cleaned$TX_DATE[i])-as.Date(heart.df.cleaned$INIT_DATE[i]))
}


}

#in this block I just use literature based variables, from the below papers:
# 1- Medved, Dennis, et al. "Improving prediction of heart transplantation outcome using deep learning techniques." Scientific reports 8.1 (2018): 3613.
# 2- Dag, Ali, et al. "A probabilistic data-driven framework for scoring the preoperative recipient-donor heart transplant survival." Decision Support Systems 86 (2016): 1-12
# I used all the variables from the first paper and FUNC_STAT_TRR, and FUNC_STAT_TCR from the second one.

# Age / AGE, AGE_DON
# Height / HGT_CM_CALC, HGT_CM_DON_CALC
# Gender / GENDER, GENDER_DON
# Weight / WGT_KG_DON_CALC, WGT_KG_TCR
# Diagnosis / DIAG
# cause of death / COD_CAD_DON
# Blood group / ABO, ABO_DON
# insulin treated diabetes / DIAB
# infection within two weeks / INFECT_IV_DRUG_TRR
# previous blood transfusion / TRANSFUSIONS
# Previously transplanted / PREV_TX
# previous cardiac surgery / CARD_SURG
# Intensive care unit / ICU
# ventiricular assist device / VAS
# Mechanical ventilation / VENTILATOR_TRR
# ECMO / ECMO
# SPP (mmHG), systolic pulmonary pressure  / HEMO_SYS_TRR
# PVR
# Duration of ischemia / ISCHTIME 
# Creatinine / CREAT_TRR
# Serum bilirubin (μmol/l) / TBILI
# PRA > 10 % / PRAMR
# HLA-DR 2 mismatch / RDR2
# Transplant era / TX_YEAR

vars<-c("AGE","AGE_DON","HGT_CM_TCR","HGT_CM_DON_CALC","GENDER","GENDER_DON","WGT_KG_DON_CALC","WGT_KG_TCR","DIAG","COD_CAD_DON", 
        "ABO","ABO_DON","DIAB","INFECT_IV_DRUG_TRR","TRANSFUSIONS","PREV_TX","CARD_SURG","ICU","VAS","VENTILATOR_TRR",
        "ECMO","HEMO_SYS_TRR","PVR","ISCHTIME","CREAT_TRR","TBILI","PRAMR_CAT10","RDR2","TX_YEAR_CAT",
        "FUNC_STAT_TRR","FUNC_STAT_TCR",
        "GSTATUS","GTIME","ID")  


# I did feature selection just based on literature
heart_df<-heart.df.cleaned[vars]

# here I made 11 consecutive dependent variables, month1, year1, year2, year3, ..., year10

  p_unit<-c(1/12,1,2,3,4,5,6,7,8,9,10)
  predict_length<-365
  
    heart_df$year0<-NA
    heart_df$year1<-NA
    heart_df$year2<-NA
    heart_df$year3<-NA
    heart_df$year4<-NA
    heart_df$year5<-NA
    heart_df$year6<-NA
    heart_df$year7<-NA
    heart_df$year8<-NA
    heart_df$year9<-NA
    heart_df$year10<-NA
  
    heart_df$GSTATUS<-as.character(heart_df$GSTATUS)
    
    heart_df<-heart_df[complete.cases(heart_df$GTIME),]
  
    for(i in 1:nrow(heart_df)){
      
    heart_df$year0[i]<-class_generator_bino(heart_df$GSTATUS[i], heart_df$GTIME[i] ,p_unit[1],predict_length)
    heart_df$year1[i]<-class_generator_bino(heart_df$GSTATUS[i], heart_df$GTIME[i] ,p_unit[2],predict_length)
    heart_df$year2[i]<-class_generator_bino(heart_df$GSTATUS[i], heart_df$GTIME[i] ,p_unit[3],predict_length)
    heart_df$year3[i]<-class_generator_bino(heart_df$GSTATUS[i], heart_df$GTIME[i] ,p_unit[4],predict_length)
    heart_df$year4[i]<-class_generator_bino(heart_df$GSTATUS[i], heart_df$GTIME[i] ,p_unit[5],predict_length)
    heart_df$year5[i]<-class_generator_bino(heart_df$GSTATUS[i], heart_df$GTIME[i] ,p_unit[6],predict_length)
    heart_df$year6[i]<-class_generator_bino(heart_df$GSTATUS[i], heart_df$GTIME[i] ,p_unit[7],predict_length)
    heart_df$year7[i]<-class_generator_bino(heart_df$GSTATUS[i], heart_df$GTIME[i] ,p_unit[8],predict_length)
    heart_df$year8[i]<-class_generator_bino(heart_df$GSTATUS[i], heart_df$GTIME[i] ,p_unit[9],predict_length)
    heart_df$year9[i]<-class_generator_bino(heart_df$GSTATUS[i], heart_df$GTIME[i] ,p_unit[10],predict_length)
    heart_df$year10[i]<-class_generator_bino(heart_df$GSTATUS[i], heart_df$GTIME[i] ,p_unit[11],predict_length)
  }

    heart_df<-heart_df[complete.cases(heart_df$year10),]
    heart_df$GSTATUS<-NULL
    heart_df$GTIME<-NULL

    # in table_cleaner_simple(Try_data,col2row_emp, ID), "Try_data" is the data we use and "col2row_emp" represent our preference of 
    # keeping columns in compare with rows, higher means more interest to keep columns than rows, ID is the ID column

    heart_max2<-table_cleaner_simple(heart_df,2,"ID")
    heart_max3<-table_cleaner_simple(heart_df,3,"ID")
    heart_max4<-table_cleaner_simple(heart_df,4,"ID")
    heart_max5<-table_cleaner_simple(heart_df,5,"ID")
    heart_max10<-table_cleaner_simple(heart_df,10,"ID")
    
# I work with heart_max10, to keep almost all the variables except for "CARD_SURG" that refers if someone had surgery, it just hass too many NAs
    heart_max<-heart_df[which(heart_df$ID %in% heart_max5$row_ID[,]),heart_max5$col_names]
    
#here save0
    table(heart_max$year0)
    table(heart_max$year1)
    table(heart_max$year2)
    table(heart_max$year3)
    table(heart_max$year4)
    table(heart_max$year5)
    table(heart_max$year6)
    table(heart_max$year7)   
    table(heart_max$year8)
    table(heart_max$year9)
    table(heart_max$year10)  


# INFECT_IV_DRUG_TRR 7000
# TRANSFUSIONS 2000
# PREV_TX 
# CARD_SURG 27000



table(heart.df.cleaned$PREV_TX)


#excluding some irrelevant variables
for(i in 1:ncol(heart_max)){
  for (j in 1:nrow(var_desc)) {
    if(names(heart_max)[i]==var_desc$VARIABLE.NAME[j]){
      if(var_desc$Hamid_TYPE[j]=="NUM"){heart_max[i]<-as.numeric(heart_max[i])
      }
    }
  }
}




```
<br />
<br />
If the levels of variables are more than 100, I assume they are numerical or there are uselessly too many levels. Some of variables are empty and some of them have a lot of levels (like states' name). Here, you can see possible values of different variables in the thoracic dataset:
```{r , message=FALSE, cache=FALSE, error=FALSE}
  DT::datatable(var_table)
```

```{r , message=FALSE, cache=FALSE, error=FALSE}

# I bring the data set who has information about which values mean NA, then convert those to NA in the dataset


 
```








