---
title: Investigating the thoracic data
author:

- Hamidreza Ahady Dolatsara^[Auburn University, [hamid@auburn.edu](mailto:hamid@auburn.edu)]
- Fadel Megahed^[University of Miami Ohio, [fmegahed@miamioh.edu](mailto:fmegahed@miamioh.edu)]
- Ying-Ju Tessa Chen^[Dayton University, [ychen4@udayton.edu](mailto:ychen4@udayton.edu)]
- Christy Evans^[Auburn University, [cje0010@auburn.edu](mailto:cje0010@auburn.edu)]
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    code_folding: hide
    df_print: paged
    number_sections: yes
    toc: yes
    toc_float: yes
  word_document:
    toc: yes
---

  

****



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In the following block, required ***R*** libraries are checked for existance and then installed if it's needed.


```{r, message=FALSE, cache=FALSE, error=FALSE}

rm(list = ls()) # clears global environment (functions and variables)
graphics.off() # clears graphics
if(!("pacman" %in% rownames(installed.packages()))){install.packages("pacman")}# pacman would be installed if it's not existed
library(pacman) 
p_load(rio)
 
```
Here we check different values that a variable might have.
```{r , message=FALSE, cache=TRUE, error=FALSE, results="hide", warning=FALSE}

thoracic_data<-import(c("https://auburn.box.com/shared/static/bcyulwlkmnhg21fmi1tvnuajv8xjhn1d.sas7bdat"))
nrows=ncol(thoracic_data)

var_table<-matrix(NA, ncol = 2, nrow = nrows)
var_table<-as.data.frame(var_table)
names(var_table) <- c("vars","data")

var_table$vars<-colnames(thoracic_data)

for(i in 1:ncol(thoracic_data)){
  if(length(as.data.frame(table(thoracic_data[i]))[,1])<100){
    var_table$data[i]<-paste(as.data.frame(table(thoracic_data[i]))[,1],collapse=",")
  }else{var_table$data[i]<-"numeric or too many levels"}
  
  
}

```
<br />
<br />
If the levels of variables are more than 100, I assume they are numerical or there are uselessly too many levels. Some of variables are empty and some of them have a lot of levels (like states' name). Here, you can see possible values of different variables in the thoracic dataset:
```{r , message=FALSE, cache=FALSE, error=FALSE}
  DT::datatable(var_table)
```



